A molecular dynamics study of the solvation of carbon dioxide and other compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-16 ionic_liquids.tex
Revision of activity coefficients section

As will be shown below, the prediction of activity coefficients requires additional simulations of the solute as a pure liquid phase, whose properties we denote here by a superscript ``$0$''.
We start by using a standard thermodynamic relation \cite{Tester} to write the solute activity coefficient $\gamma_s$ in terms of chemical potentials as
\begin{equation}
\label{eq:gamma}
\ln \gamma_s = \frac{\mu_s^l - \mu_s^{l,0}}{R T} - \ln x_s,
\end{equation}

Next, we replace $\mu_s^l$ and $\mu_s^{l,0}$ by their corresponding instances of Eq.~\eqref{eq:solute chemical potential}. In this step, we note that $c_s = \tilde{\rho} x_s$ and $c_s^0 = \tilde{\rho}_s^0$, where $\tilde{\rho}$ and $\tilde{\rho}_s^0$ are the molar densities of the solution and of pure liquid solute, respectively. Thus, the expression we obtain for computing an activity coefficient from two independent simulations is
\begin{equation}
\ln \gamma_s = \frac{\Delta G_\text{sim} - \Delta G_\text{sim}^{0}}{R T} + \ln \frac{\tilde{\rho}}{\tilde{\rho}_s^0}.
\end{equation}

@@

Having selected the model of \ce{[emim][B(CN)_4]} that best reproduces the experimental $K_s$ of \ce{CO_2}, we next calculate the infinite-dilution activity coefficients ($\gamma^{\, \infty}$) of benzene, hexane, cyclohexane, ethanol and water in \ce{[emim][B(CN)_4]}.

The task of accurately predicting $\gamma^{\, \infty}$ is challenging since, according to Eq.~\eqref{eq:gamma_inf}, one should reproduce the solvation free energies in both the pure solute and the solvent.
In Table~\ref{table:mu_solutes} we show the simulated densities and the solvation free energies $(\Delta G^{0}_\text{sim})$ of the pure solutes, which we compare against the experimental data extracted from the work of Chang~\cite{Chang_2009}.
It should be pointed out that we were unable to find the experimental $\Delta G^{0}_\text{exp}$ for cyclohexane.
Note that only in the case of ethanol we successfully predicted the $\Delta G^{0}_\text{exp}$.
The relative errors for the other solutes are around 10\%.

The results of $\gamma^{\, \infty}$ are presented in Table~\ref{table:gamma}.
For benzene and water, the predictions are very accurate, but this may be due to fortuitous error compensations given the divergences shown in Table~\ref{table:mu_solutes}.
From Table \ref{table:gamma}, one sees that the experimental $\gamma^{\, \infty}$ for hexane at $298.15~\text{K}$ and $303~\text{K}$ considerably differ.
It is therefore difficult to assess the reliability of our results in this case.

\begin{table*}
	\centering
	\caption{Solvation free energy at $T = 298.15~\text{K}$ for the pure solutes benzene, hexane, cyclohexane, ethanol and water.}
	\begin{adjustbox}{max width=\textwidth}
		\begin{tabular}{ccccccccc}
			\hline\hline
			Solute & $\rho_\text{sim}$ & error & $\Delta G_\text{vdW}$  & $\Delta G_\text{Coul}$  & $\Delta G^0_\text{sim}$ & $\Delta G^0_\text{exp}$   & error \\
			& (g/cm\textsuperscript{3}) & (\%) & (kcal/mol) &  (kcal/mol) &  (kcal/mol)   & (kcal/mol)   & (\%) \\
			\hline
			hexane & $0.6513 \pm 0.0003$ & $0.5$ & $-3.67  \pm  0.03$ & $0.011 \pm 0.007$ & $-3.66 \pm 0.03$ & $-4.06$ & $9.90$ \\
			cyclohexane & $0.7692 \pm 0.0007$ & $0.6$ & $-4.39 \pm 0.01$ & $0$ & $-4.39 \pm 0.01$ & n/a & n/a  \\
			benzene & $0.8611 \pm 0.0002$ & $1.4$ & $-3.75  \pm 0.02$ & $-0.381 \pm 0.007$ & $-4.14 \pm 0.02$ & $-4.56$ & $9.32$  \\
			ethanol & $0.781 \pm 0.008$ & $1.01$  & $-0.93 \pm 0.01$ & $-4.15 \pm 0.02$  & $-5.08  \pm 0.02$  & $-5.08$ & $0$ \\
			water & $0.9948 \pm 0.0003$ & $0.2$ & $2.015 \pm 0.004$ & $-9.01 \pm 0.02$ & $-6.99 \pm 0.02$ & $-6.33$  & $10.43$ \\
			\hline\hline
			\label{table:mu_solutes} 
		\end{tabular}
	\end{adjustbox}
\end{table*}

>>>>

As will be shown below, the prediction of activity coefficients requires additional simulations of the solute as a pure liquid phase, whose properties we denote here by a superscript ``$0$''.
We start by using a standard thermodynamic relation \cite{Tester} to write the solute activity coefficient $\gamma_s$ in terms of chemical potentials as
\begin{equation}
\ln \gamma_s = \frac{\mu_s^l - \mu_s^{l,0}}{R T} - \ln x_s,
\end{equation}

Next, we replace $\mu_s^l$ and $\mu_s^{l,0}$ by their corresponding instances of Eq.~\eqref{eq:solute chemical potential}. In this step, we note that $c_s = \tilde{\rho} x_s$ and $c_s^0 = \tilde{\rho}_s^0$, where $\tilde{\rho}$ and $\tilde{\rho}_s^0$ are the molar densities of the solution and of pure liquid solute, respectively. Thus, the expression we obtain for computing an activity coefficient from two independent simulations is
\begin{equation}
\label{eq:gamma}
\ln \gamma_s = \frac{\Delta G_\text{sim} - \Delta G_\text{sim}^{0}}{R T} + \ln \frac{\tilde{\rho}}{\tilde{\rho}_s^0}.
\end{equation}

@@

Having selected the model of \ce{[emim][B(CN)_4]} that best reproduces the experimental Henry's law constant of \ce{CO_2}, we extend its use to the calculation of infinite-dilution activity coefficients ($\gamma^\infty$) of other solutes in \ce{[emim][B(CN)_4]}.
Due to the availability of experimental data \cite{Yan_2010, Doma_ska_2011}, the solutes chosen for this study are benzene, hexane, cyclohexane, ethanol, and water.

The task of accurately predicting $\gamma^\infty$ is challenging because, according to Eq.~\eqref{eq:gamma}, one needs to reproduce the solvation free energies of a solute molecule in both the solvent and the solute itself as a pure liquid.
In Table~\ref{table:mu_solutes}, we show the simulated densities ($\rho^0_s$) and solvation free energies $(\Delta G^{0}_\text{sim})$ of the pure solutes, which we compare against experimental data extracted from the work of Chang \cite{Chang_2009}.
It is worth remarking that we were unable to find the experimental $\Delta G^{0}_\text{exp}$ for cyclohexane.
Note that only in the case of ethanol we successfully predicted the $\Delta G^{0}_\text{exp}$.
The relative errors for the other solutes are around 10\%.

\begin{table*}
	\centering
	\caption{Solvation free energy at $T = 298.15~\text{K}$ for the pure solutes benzene, hexane, cyclohexane, ethanol and water.}
	\begin{adjustbox}{max width=\textwidth}
		\begin{tabular}{ccccccccc}
			\hline\hline
			Solute & $\rho^0_s$ & error & $\Delta G_\text{vdW}$  & $\Delta G_\text{Coul}$  & $\Delta G^0_\text{sim}$ & $\Delta G^0_\text{exp}$   & error \\
			& (g/cm\textsuperscript{3}) & (\%) & (kcal/mol) &  (kcal/mol) &  (kcal/mol)   & (kcal/mol)   & (\%) \\
			\hline
			hexane & $0.6513 \pm 0.0003$ & $0.5$ & $-3.67  \pm  0.03$ & $0.011 \pm 0.007$ & $-3.66 \pm 0.03$ & $-4.06$ & $9.90$ \\
			cyclohexane & $0.7692 \pm 0.0007$ & $0.6$ & $-4.39 \pm 0.01$ & $0$ & $-4.39 \pm 0.01$ & n/a & n/a  \\
			benzene & $0.8611 \pm 0.0002$ & $1.4$ & $-3.75  \pm 0.02$ & $-0.381 \pm 0.007$ & $-4.14 \pm 0.02$ & $-4.56$ & $9.32$  \\
			ethanol & $0.781 \pm 0.008$ & $1.01$  & $-0.93 \pm 0.01$ & $-4.15 \pm 0.02$  & $-5.08  \pm 0.02$  & $-5.08$ & $0$ \\
			water & $0.9948 \pm 0.0003$ & $0.2$ & $2.015 \pm 0.004$ & $-9.01 \pm 0.02$ & $-6.99 \pm 0.02$ & $-6.33$  & $10.43$ \\
			\hline\hline
			\label{table:mu_solutes} 
		\end{tabular}
	\end{adjustbox}
\end{table*}

The results concerning the solvation free energies of all solutes in the ionic liquid are presented in Table~\ref{table:gamma}, as well as the computed infinite-dilution activity coefficients.
For benzene and water, the predictions are very accurate, but this is certainly due to fortuitous error compensations, given the discrepancies observed in Table~\ref{table:mu_solutes}.
In the case of hexane, one can see in Table \ref{table:gamma} that the available experimental values of $\gamma^\infty$ in \ce{[emim][B(CN)_4]} at very close temperatures ($298.15~\text{K}$ and $303~\text{K}$) differ considerably.
It is, therefore, difficult to assess the reliability of our results in this case.
